\begin{longtable}{TD}
\caption{Uncertainty, confidence, residual, and risk terms.}
\label{tab:terms-uncertainty-confidence-risk}\\

\toprule
Term & Definition \\
\midrule
\endfirsthead

\toprule
Term & Definition \\
\midrule
\endhead

\bottomrule
\endfoot

\termrow{uncertainty}{
Unreliability or unpredictability of a task behavior estimate. Uncertainty may
come from missing features, hidden state, measurement error, drift, heavy tails,
multimodality, or insufficient history.
}

\termrow{entropy}{
A possible measure or family of measures for unpredictability. The term should
only be used after specifying what distribution, signal, or behavior process is
being measured.
}

\termrow{confidence}{
The degree of trust in an estimate, behavior profile, graph edge, feature
mapping, or model assumption. Confidence is distinct from uncertainty.
}

\termrow{residual error}{
The part of observed behavior not explained by the current model estimate.
Residual error exposes missing features, hidden influences, drift, or model
mismatch.
}

\termrow{prediction error}{
The difference between predicted and observed value for a resource, occupancy,
duration, outcome, or risk-relevant dimension.
}

\termrow{drift}{
A change in task behavior, feature mapping, graph relationship, context, or
profile over time. Drift is a source of uncertainty and reduced confidence.
}

\termrow{variance}{
The statistical spread of observed values. Variance is a variability signal but
is not identical to entropy.
}

\termrow{tail ratio}{
A ratio comparing a tail quantile with a central quantile, such as p95/p50 or
p99/p50. Tail ratios help identify high-tail-risk behavior.
}

\termrow{multimodality}{
The presence of multiple behavioral regimes or modes in the observed behavior
of a task kind or task family.
}

\termrow{operational risk}{
The consequence-weighted danger of underestimating, misclassifying, or
mishandling a task attempt. Operational risk is separate from task cost.
}

\termrow{consequence}{
The severity of the effect caused by an incorrect estimate or control decision,
such as delayed processing, queue starvation, SLO violation, replication lag, or
cascading overload.
}

\termrow{blast radius}{
The scope affected by an error or overload event, such as a single task, queue,
worker pool, tenant, replica set, dependency, cluster, or control plane.
}

\termrow{SLO violation risk}{
The operational risk that a task attempt, task kind, queue, tenant, or service
will violate a service-level objective.
}

\termrow{queue starvation risk}{
The risk that long-running or high-occupancy tasks delay or prevent other tasks
from receiving execution capacity.
}

\termrow{retry amplification risk}{
The risk that retries increase work, dependency pressure, queue occupancy, or
worker occupancy enough to worsen overload.
}

\termrow{cascade overload risk}{
The risk that local overload propagates to other queues, workers, dependencies,
tenants, or control loops.
}

\end{longtable}
