\begin{longtable}{TD}
\caption{Manifest, registry, and domain-pack terms.}
\label{tab:terms-manifests-registries-domain-packs}\\

\toprule
Term & Definition \\
\midrule
\endfirsthead

\toprule
Term & Definition \\
\midrule
\endhead

\bottomrule
\endfoot

\termrow{Task Behavior Manifest}{
A declarative description of the behavior model for a task kind or task family.
It may describe features, domains, graph nodes, graph edges, observation
requirements, profiling policy defaults, known gaps, and versioning.
}

\termrow{manifest}{
A machine-readable or human-readable specification used to declare part of the
task behavior model.
}

\termrow{domain pack}{
An extension package that provides domain-specific metrics, feature mappings,
feature descriptors, graph templates, examples, or profiling defaults.
}

\termrow{metric descriptor registry}{
A registry describing raw metrics and how they can be interpreted, aggregated,
and mapped to canonical features.
}

\termrow{graph declaration}{
The part of a manifest that declares graph nodes, graph edges, influence roles,
sensitivities, confidence, and applicability conditions.
}

\termrow{profiling policy declaration}{
The part of a manifest that declares default observation strategies, required
observation points, sampling rules, and profiling overhead limits.
}

\termrow{schema}{
A machine-validation structure for manifests, descriptors, registries, domain
packs, or graph declarations.
}

\termrow{versioning}{
The practice of assigning versions to schemas, manifests, domain packs, task
kinds, feature registries, or graph structures in order to manage evolution and
compatibility.
}

\termrow{compatibility}{
The ability of manifests, schemas, domain packs, task versions, and graph
structures to work together without semantic or validation conflicts.
}

\termrow{coverage}{
The set of task kinds, domains, features, metrics, observation points, or risks
covered by a model, manifest, or domain pack. Coverage is not the same as global
completeness.
}

\termrow{known gap}{
A missing, unavailable, weak, or intentionally unsupported feature, observation,
domain, or graph relationship. Known gaps should remain visible rather than
being hidden inside the model.
}

\termrow{residual visibility}{
The ability of the model to expose unexplained behavior through residual error,
reduced confidence, or missing-feature indicators.
}

\termrow{canonical namespace}{
A stable naming scheme for canonical features, resource dimensions, occupancy
dimensions, and risk dimensions, such as \texttt{context.network.instability}
or \texttt{occupancy.worker.slot\_time}.
}

\end{longtable}
