\begin{longtable}{TD}
\caption{Core terms used in the main model.}
\label{tab:core-terms}\\

\toprule
Term & Definition \\
\midrule
\endfirsthead

\toprule
Term & Definition \\
\midrule
\endhead

\bottomrule
\endfoot

\termrow{task}{A generic unit of work handled by a queue-worker system.}

\termrow{task kind}{
The semantic type of work submitted to a queue. A task kind selects reusable
model structure, relevant features, cost dimensions, graph fragments, and
profiling-policy priors, but it does not determine one universal cost value.
}

\termrow{task instance}{
A concrete logical unit of work placed into a queue. A task instance belongs to
a task kind and may produce one or more attempts.
}

\termrow{task attempt}{
One concrete execution of a task instance. Resource usage, worker occupancy,
observations, outcomes, and residuals are estimated or observed at attempt
granularity.
}

\termrow{worker}{
An executor that reserves, executes, retries, acknowledges, discards, or fails
task attempts.
}

\termrow{task behavior}{
The broader profile of a task attempt or kind, including cost estimates,
symptoms, outcomes, reliability diagnostics, risk-relevant effects, and
policy-relevant triggers.
}

\termrow{execution context}{
The input-side environment and state under which a task attempt executes. It may
influence cost, symptoms, outcomes, diagnostics, and risk, but it is not itself
a cost output.
}

\termrow{observation}{
Any measurement, event, trace, profile sample, counter, log entry, or
domain-specific signal exposed by a system component. An observation becomes a
model variable only after it is mapped to a semantic role.
}

\termrow{canonical feature}{
A typed and normalized model input derived from task metadata, execution
context, raw observations, or domain-specific signals. Canonical features form
the input vector available to estimators.
}

\termrow{influence domain}{
A semantic role assigned to a model variable. Influence domains separate task
identity, intrinsic work, execution context, resource demand, worker occupancy,
symptoms, outcomes, reliability diagnostics, operational risk, and
profiling-policy triggers.
}

\termrow{cost dimension}{
A typed dimension of attempt-attributed cost with role, unit, value domain,
attribution boundary, observation semantics, and composition semantics.
}

\termrow{estimate object}{
A dimension-level estimate that may contain a point value, quantiles,
distributional summaries, conservative envelopes, missingness information, or
other diagnostic material.
}

\termrow{resource demand}{
The modeled active resource need expected for a task attempt along a resource
dimension. It is distinct from observed resource usage and environment-side
pressure.
}

\termrow{resource usage}{
The observed attempt-attributed use of a resource during execution.
}

\termrow{worker occupancy}{
Bounded execution capacity held by a task attempt over time. Occupancy exists
only during intervals in which the capacity is assigned to the attempt.
}

\termrow{worker-slot time}{
The time during which a task attempt holds a worker execution slot.
}

\termrow{symptom}{
An intermediate observed signal that indicates abnormal or relevant behavior but
is not itself a final outcome, a cost dimension, or a root-cause claim.
}

\termrow{outcome}{
The completion or lifecycle result of a task attempt, such as success, failure,
timeout, cancellation, retryable error, dead-lettering, or lease expiration.
}

\termrow{reliability diagnostic}{
A diagnostic that describes the reliability of a cost estimate or behavior
profile, such as residual magnitude, missingness, low sample support, drift,
tail ratio, multimodality, or confidence.
}

\termrow{uncertainty}{
Unreliability or unpredictability of a task behavior estimate. In this paper,
uncertainty is represented through reliability diagnostics rather than treated
as a cost dimension.
}

\termrow{operational risk}{
The consequence-weighted danger of underestimating, misclassifying, or
mishandling a task attempt. Risk is downstream from cost, diagnostics, outcomes,
and consequence information.
}

\termrow{profiling-policy trigger}{
A model variable or decision signal that selects or justifies an observation
strategy, such as minimal accounting, detailed accounting, tracing, tail-aware
profiling, or diagnostic profiling.
}

\termrow{Task Behavior Graph}{
A typed directed graph associated with a task kind or family. It connects
canonical features and derived model variables across influence domains and
supports estimation of cost, diagnostics, risk, and profiling-policy triggers.
}

\termrow{profiling policy}{
The selected observation and instrumentation strategy for a task kind, attempt,
or execution context.
}

\end{longtable}
