\section{Core Terminology and Notation}
\label{sec:terminology-notation}

This section defines only the core terms required to state the problem and
introduce the model. The paper uses a larger terminology set, but the complete
reference glossary is deferred to Appendix~\ref{app:extended-terminology} so
that the main text can proceed from the problem statement to the model without
front-loading implementation and lifecycle details.

The terminology is role-specific. Task identity and intrinsic task features
describe what work is requested. Execution context describes the environment and
state under which an attempt executes. Resource demand and worker occupancy are
cost-side outputs attributed to the attempt. Observations provide evidence;
canonical features are typed model inputs derived from observations or metadata.
Uncertainty and operational risk are downstream interpretations of estimates,
residuals, missing information, and consequences. These roles are kept distinct
because the same operational event can connect several roles without making them
synonyms. For example, degraded dependency health is an execution-context
factor; dependency wait time is an observed symptom or intermediate duration;
connection hold time and worker-slot time are occupancy dimensions; and queue
starvation is a possible operational risk.

\subsection{Core Terms}
\label{sec:core-terms}
\input{tables/terminology/core-terms}

\subsection{Notation}
\label{sec:terminology-symbols}
\input{tables/terminology/12-notation}
