\section{Problem Statement}
\label{sec:problem-statement}

The preceding section motivates a gap between available observations and the
model needed to profile logical task attempts in queue-worker systems. This
section states the problem addressed by the paper. It defines the system model,
the information available to the profiler, the outputs expected from the
profiling model, the main constraints, and the non-goals. The purpose is to
separate the problem from the model proposed in later sections.

At a high level, the problem is to estimate the behavior of a task attempt from
its task kind, task-instance features, execution-context observations, and
available history, while respecting profiling overhead. The estimate must not be
reduced to a single scalar cost. It must preserve active resource demand,
worker occupancy, uncertainty, and operational risk as distinct outputs because
these quantities support different operational decisions.

\subsection{System Model}
\label{sec:problem-system-model}

We consider queue-worker systems in which producers submit task instances to one
or more queues and workers execute task attempts. A task kind identifies the
semantic type of work requested. A task instance is one queued unit of that kind.
A task attempt is one concrete execution of a task instance. A single task
instance may produce multiple attempts because of retries, lease expiration,
timeouts, cancellation, worker failure, or rescheduling.

The profiling boundary is the task attempt. This boundary is narrower than a
worker process, container, node, or worker pool, but broader than a single
function call or trace span. An attempt may include multiple phases, such as
input fetching, parsing, computation, dependency calls, retries, output writes,
commits, acknowledgement, and cleanup. These phases may be sequential,
parallel, conditional, retried, or partially overlapping. The paper does not
assume that every phase is perfectly instrumented.

Each attempt executes under an execution context. The context may include worker
state, node pressure, dependency health, network condition, disk pressure, cache
state, replica state, tenant pressure, retry state, and domain-specific hidden
state. Some context factors are directly observable, some are derived from raw
observations, and some remain hidden. Hidden or missing factors are represented
indirectly through residual error, uncertainty, and reduced confidence.

The model is task-kind-specific. Different task kinds may require different
features, different observation points, and different profiling policies. A
replication task kind may be sensitive to network condition, replica health,
replication lag, disk pressure, and retry state. A compute-oriented transform
may be more sensitive to record count, algorithm mode, CPU pressure, allocation
behavior, and memory pressure. The system therefore cannot assume one global
feature set or one global metric list that is equally relevant to all task
kinds.

\subsection{Available Information}
\label{sec:problem-available-information}

For an attempt $a$, the profiler may have access to four classes of
information. First, task identity describes what is being executed. It may
include the task kind, task version, queue, producer, tenant, tenant class,
priority, deadline, and attempt number. Task identity is necessary for selecting
a task-kind-specific model, but it is not by itself a cost estimate.

Second, intrinsic task features describe work contained in the task instance.
Examples include payload size, record count, object count, partition size,
batch size, input format, operation mode, declared resource hints, or domain
metadata. These features may be available before execution and can provide a
baseline for expected work. Their presence and precision are domain-specific.

Third, execution-context observations describe the environment in which the
attempt is executed or is expected to execute. Examples include worker-pool
pressure, node CPU or memory pressure, disk queue depth, network instability,
dependency health, connection-pool pressure, cache warmth, replica health,
replication lag, tenant quota usage, and retry state. These observations may be
fresh, stale, missing, aggregated, or only indirectly related to the attempt.

Fourth, historical observations describe previous attempts of the same or
related task kinds. They may include observed resource usage, worker occupancy,
durations, retry counts, outcomes, residual errors, tail quantiles, or drift
indicators. Historical data may be unavailable for new task kinds, new task
versions, rare contexts, or recently changed workloads. The problem must
therefore support both calibrated and cold-start cases.

Raw observations are not assumed to be model-ready. Metrics, logs, traces,
runtime profiles, and domain events must first be interpreted, normalized, and
mapped into canonical features. This mapping is part of the profiling problem
because the same raw metric can play different roles in different domains, and
similar operational influences can be exposed through different raw metrics.

\subsection{Desired Outputs}
\label{sec:problem-desired-outputs}

The profiler should produce a structured estimate for a task attempt or for a
class of comparable attempts. The first output is an estimate of active resource
demand. For each relevant resource dimension $r \in \mathcal{R}$, the model
should estimate $\widehat{D}_r(a)$ or a distribution over possible values.
Relevant dimensions may include CPU time, memory peak, disk reads, disk writes,
network input, network output, allocation volume, or accelerator usage.

The second output is an estimate of worker occupancy. For each relevant
occupancy dimension $o \in \mathcal{O}$, the model should estimate
$\widehat{U}_o(a)$ or a distribution over possible values. Relevant dimensions
may include worker-slot time, lease time, connection hold time, lock hold time,
semaphore hold time, and tenant concurrency time. Occupancy is modeled
separately because an attempt can hold scarce execution capacity without
consuming much active CPU or memory.

The third output is uncertainty and confidence. The model should indicate how
reliable its estimates are, whether residual error is expected to be high,
whether behavior has drifted, whether historical samples are insufficient, and
whether tail behavior is operationally significant. The paper does not require a
single universal uncertainty metric. Later sections may use residuals,
confidence values, quantile ratios, drift indicators, or other measures where
appropriate.

The fourth output is operational risk. Risk is not the same as task cost. It
captures the consequence of underestimating, misclassifying, or mishandling an
attempt. Relevant risks may include queue starvation, retry amplification,
SLO violation, replication lag, tenant impact, dependency overload, and
cascading overload. Risk depends not only on estimated demand but also on the
blast radius and the operational meaning of the task kind.

The fifth output is a profiling-policy recommendation. Given the task kind,
context, uncertainty, risk, and profiling budget, the system should determine
what observation strategy is justified. Possible strategies include minimal
accounting, detailed accounting, sampled tracing, tail-aware profiling,
diagnostic profiling, or isolated diagnostic execution. The policy output is a
profiling decision, not a final scheduling or routing decision.

\subsection{Core Challenges}
\label{sec:problem-core-challenges}

The first challenge is conditional behavior. Attempts with the same task kind
and similar intrinsic features may behave differently under different execution
contexts. A model based only on task kind, queue, or payload bucket can provide a
baseline, but it cannot explain context-dependent amplification, suppression,
or drift.

The second challenge is multidimensionality. Active resource demand, bounded
worker occupancy, execution-context pressure, uncertainty, and operational risk
have different meanings and units. Combining them prematurely into one scalar
hides information needed by downstream control layers. A scalar score may be a
useful projection, but it should be derived from a structured estimate rather
than replace it.

The third challenge is attribution. Process-level, container-level, node-level,
and queue-level observations are often easier to collect than task-attempt-level
observations. These aggregate observations may indicate pressure, but they do
not automatically explain which task kind, instance, attempt, phase, context
factor, or dependency produced it.

The fourth challenge is semantic interpretation. Raw observations do not state
whether they represent intrinsic work, execution context, active resource usage,
worker occupancy, a symptom, an outcome, an uncertainty indicator, or a risk
indicator. Without explicit semantic roles, a model may double-count related
signals. For example, network instability, retry count, wall-clock duration,
and worker-slot time may all reflect one underlying degradation, but they should
not be treated as independent causes.

The fifth challenge is compositional behavior. A task kind may include multiple
behavior components, such as CPU processing, network transfer, dependency calls,
retries, commits, and cleanup. Some quantities compose by summation, some by
maximum, some by critical path, and some by retry expansion or
occupancy-specific rules. The profiling problem must therefore preserve enough
structure to combine observations consistently.

The sixth challenge is partial observability. Important influences may be
missing, stale, hidden, or domain-specific. The model cannot assume that all
relevant state is observable. Instead, unexplained behavior should remain
visible through residual error, reduced confidence, or a higher profiling level.

The seventh challenge is profiling overhead. Instrumentation is not free.
Detailed tracing, runtime profiling, high-cardinality labels, frequent
sampling, and telemetry export can affect the workload being measured. A
profiling model must therefore decide not only what would be useful to observe,
but what is worth observing under an overhead budget.

\subsection{Scope and Assumptions}
\label{sec:problem-scope-assumptions}

This paper assumes that task attempts can be identified at least by task kind
and attempt boundary. The model is less useful when all work is invisible inside
one undifferentiated worker loop. The paper does not require perfect
instrumentation, but it assumes that minimal accounting is possible for at least
some attempts, including start, finish, outcome, and task identity.

The paper assumes that some task metadata or intrinsic features may be available
before or during execution. The exact features are domain-specific. A small
validation task, a replication task, an external API fan-out task, and a media
transformation task need not expose the same feature set.

The paper assumes that execution-context observations may be available but
incomplete. These observations may come from workers, nodes, queues,
dependencies, telemetry systems, domain services, or historical profiles. The
model must tolerate missing, stale, noisy, or partially aggregated context.

The paper assumes that profiling decisions are useful even when final
scheduling, routing, placement, or admission decisions are made by another
layer. The profiling model supplies structured estimates and observation
policies. Downstream control layers may consume these outputs, but their full
design is outside the scope of the paper.

The paper also assumes an open-world metric environment. No core system can know
all possible domain-specific metrics in advance. The model must therefore allow
raw observations to be mapped into canonical features and must keep missing or
unexplained influences explicit instead of hiding them behind a fixed global
metric list.

\subsection{Non-Goals}
\label{sec:problem-non-goals}

This paper does not propose a complete distributed scheduler. It does not define
a placement algorithm, routing algorithm, gossip protocol, load-balancing policy,
or cluster-wide resource allocator. Such systems may use the estimates produced
by task profiling, but they are not the object of this paper.

This paper does not replace metrics, logs, traces, runtime profiles, or existing
observability tools. These systems provide observations. The problem addressed
here is how to interpret observations for task-kind-specific profiling and how
to decide which observations are justified.

This paper does not claim that Task Behavior Graphs prove causality. A graph can
make modeled influence assumptions explicit, but causal validation requires
additional evidence, experiments, interventions, or domain knowledge. The graph
is a profiling model, not a causal proof.

This paper does not attempt to define a globally complete taxonomy of all
workload metrics. It assumes that new domains can introduce new observations,
features, and behavior components. Completeness is task-kind- and domain-specific
rather than global.

This paper does not require exact prediction for arbitrary tasks. The goal is to
produce structured estimates, expose uncertainty, and guide profiling policy.
High residual error, drift, missing features, or high tail risk are valid model
outputs, not merely failures to hide.

Finally, this paper does not specify a full production implementation of a
package manager, expression language, runtime virtual machine, or cross-language
manifest compiler. Later sections discuss manifests and implementation
considerations only to the extent needed to make the proposed model concrete.
