\section{Task Cost Model}
\label{sec:task-cost-model}

Section~\ref{sec:problem-statement} states the profiling problem as a structured
estimation problem. The profiler may observe task identity, intrinsic task
features, execution context, task-kind model state, and historical behavior, and
it should produce cost estimates, reliability diagnostics, risk interpretation,
and profiling-policy recommendations. This section defines only the first of
those output classes: attempt-attributed cost estimates.

The cost model is intentionally narrower than the complete task-behavior model.
It does not define influence domains, graph topology, feature mapping,
uncertainty metrics, risk scoring, or profiling-policy selection. Those layers
consume or organize cost information, but they are not part of the cost object
itself. The purpose of this section is to define what a cost estimate is, what it
is estimated over, how it is typed, how it is related to observed values, and why
a single scalar load value is insufficient.

The section proceeds in the order in which a reader needs the concepts. First,
it defines the attempt-level cost object. Second, it separates active resource
demand from worker occupancy. Third, it defines the feature inputs and
task-kind-specific model state used by estimators. Fourth, it introduces typed
cost dimensions and dimension-level estimate objects. Fifth, it defines
dimension-specific estimators and the common baseline--amplification--overhead
decomposition. Finally, it explains overhead, observations, residuals,
distributional estimates, and the boundary between the cost model and the later
graph, diagnostic, risk, and policy layers.

\subsection{Attempt-Level Cost Object}
\label{sec:cost-attempt-level}

The unit of cost attribution is the task attempt. A task kind defines reusable
model structure: which features are relevant, which cost dimensions are
estimated, and which estimators or graph fragments may be used. A task instance
supplies the logical work to be performed. A task attempt supplies the concrete
execution under one execution context. Therefore, structure is task-kind-level,
but evaluation is attempt-level.

Let $a \in \mathcal{A}$ be a task attempt. The attempt executes task instance
$i(a) \in \mathcal{I}$, and the task kind of that instance is
$k(i(a)) \in \mathcal{K}$. For readability, write
\[
k_a = k(i(a)).
\]
The symbol $k_a$ does not introduce a new entity. It is only a shorthand for the
kind of the task instance executed by attempt $a$. This shorthand is useful
because all task-kind-specific objects used below, such as relevant dimensions
and estimator state, are selected by $k_a$.

The attempt-level cost estimate is a structured object:
\[
\widehat{\mathbf{C}}(a)
=
\left(
\widehat{\mathbf{D}}(a),
\widehat{\mathbf{U}}(a)
\right).
\]
The first component, $\widehat{\mathbf{D}}(a)$, contains estimates for active
resource-demand dimensions. The second component, $\widehat{\mathbf{U}}(a)$,
contains estimates for worker-occupancy dimensions. The hat indicates that this
is a model estimate rather than an observed value. The bold symbols indicate
that the estimate is vector-valued or structured rather than scalar.

For task kind $k$, let $\mathcal{R}_k$ be the set of resource-demand dimensions
that the model estimates for that kind, and let $\mathcal{O}_k$ be the set of
worker-occupancy dimensions that the model estimates for that kind. For attempt
$a$, the two components are
\[
\widehat{\mathbf{D}}(a)
=
\left\{
\widehat{E}_r(a)
\mid
r \in \mathcal{R}_{k_a}
\right\},
\qquad
\widehat{\mathbf{U}}(a)
=
\left\{
\widehat{E}_o(a)
\mid
o \in \mathcal{O}_{k_a}
\right\}.
\]
The elements of these sets are estimate objects, not only numeric point values.
This distinction matters because the model may need to preserve a point
estimate, tail quantiles, conservative envelopes, missingness information, and
other diagnostic material for the same cost dimension. A downstream policy can
choose which projection of an estimate object it needs, but the cost model
should not discard the rest of the information prematurely.

This representation also explains why task cost is not modeled as a universal
scalar. CPU seconds, memory bytes, network bytes, slot-seconds, lock-seconds,
and connection-seconds have different units and different composition rules. A
single score can be derived later by a policy, for example
\[
s_\rho(a)
=
\pi_\rho\!\left(
\widehat{\mathbf{D}}(a),
\widehat{\mathbf{U}}(a)
\right),
\]
where $\pi_\rho$ is a policy-specific projection. However, $s_\rho(a)$ is not
the cost model. It is a decision-specific view of the structured cost estimate.

As a running example, consider a replication task kind. The task kind may define
resource-demand dimensions such as network output bytes and disk read bytes, and
worker-occupancy dimensions such as worker-slot time and connection hold time.
Two attempts of this kind may use the same model structure, but they can produce
different estimate objects because one attempt runs with healthy peers and a
small payload while another runs with degraded peers, packet loss, and retry
state. The task kind selects the model; the attempt determines the evaluated
cost.

\subsection{Resource Demand and Worker Occupancy}
\label{sec:cost-demand-occupancy}

The cost object separates two families because they answer different questions.
Resource demand asks how much active resource use is expected or observed for
the attempt. Worker occupancy asks how much bounded execution capacity is held
by the attempt over time. These quantities may be correlated, but they are not
interchangeable.

A resource-demand dimension describes active resource work attributable to the
attempt. Examples include CPU time, allocation volume, memory peak, disk read
bytes, disk write bytes, network input bytes, network output bytes, or
accelerator time. For a resource dimension $r \in \mathcal{R}_{k_a}$, the
estimate object is $\widehat{E}_r(a)$, and its point projection is
\[
\widehat{Y}_r(a)
=
\operatorname{point}_r\!\left(\widehat{E}_r(a)\right).
\]
The interpretation of $\widehat{Y}_r(a)$ depends on the descriptor of
dimension $r$. For CPU time, it may represent attributed CPU seconds. For
network output, it may represent attributed bytes. For memory peak, it may
represent a maximum rather than a sum. Therefore, a resource-demand dimension is
not defined only by a metric name; it is defined by its descriptor, attribution
boundary, and composition semantics.

Worker occupancy describes bounded capacity held by an attempt. Examples
include worker-slot time, lease time, connection hold time, lock hold time,
semaphore-permit time, and tenant-concurrency time. For an occupancy dimension
$o \in \mathcal{O}_{k_a}$, the estimate object is $\widehat{E}_o(a)$, and its
point projection is
\[
\widehat{Y}_o(a)
=
\operatorname{point}_o\!\left(\widehat{E}_o(a)\right).
\]
The unit of an occupancy dimension is usually a capacity-time unit, such as
slot-seconds, connection-seconds, lock-seconds, or permit-seconds.

The interval semantics of occupancy can be written explicitly. Let
$\mathcal{T}_{o,a}$ be the set of time intervals during which the capacity
associated with occupancy dimension $o$ is held by attempt $a$. If an attempt
can hold more than one unit of that capacity, let $m_o(t,a)$ denote the amount
of capacity held at time $t$. When the required interval observations are
available, the observed occupancy value can be understood as
\[
Y_o(a)
=
\int_{\bigcup \mathcal{T}_{o,a}}
m_o(t,a)\,dt,
\qquad
o \in \mathcal{O}_{k_a}.
\]
For a single worker slot, $m_o(t,a)$ is usually $1$ while the slot is held and
$0$ otherwise. For semaphore permits, $m_o(t,a)$ may be the number of permits
held. For dependency connections, it may be the number of connections held by
the attempt. The formula is not a required implementation. It states the model
semantics: occupancy is capacity held over time, not wall-clock duration by
itself.

This distinction avoids a common error. Suppose an attempt waits on a degraded
dependency for thirty seconds. The dependency degradation is execution context.
The dependency wait is an observed waiting signal or symptom. The attempt's
wall-clock duration is an elapsed-time outcome. Worker-slot time increases only
if the worker slot remains held during the wait. Connection hold time increases
only if a dependency connection remains held during the wait. Thus, the same
incident may affect several variables, but only some of them are cost
dimensions.

The separation also explains why active resource demand alone is insufficient.
A CPU-heavy transformation may use substantial CPU time and allocation volume
while holding a worker slot for a short interval. A dependency-bound task may
use little CPU while holding a worker slot or connection for a long interval.
Both attempts can be costly, but they consume different kinds of capacity and
should be represented by different dimensions.

\subsection{Feature Projections and Model State}
\label{sec:cost-feature-projections-model-state}

The cost model is evaluated from attempt-level input features and reusable
task-kind state. Let $z_a \in \mathcal{Z}_{k_a}$ be the canonical input-feature
vector available for attempt $a$ after feature mapping and normalization. The
vector may be partial and may include explicit missingness indicators. It
contains information available to the estimator before or during the attempt. It
does not contain post-execution observed output values for the same attempt.

The canonical input vector is separated into role-specific projections:
\[
\mathbf{x}_a = P_X^{k_a}(z_a),
\qquad
\mathbf{c}_a = P_C^{k_a}(z_a).
\]
The projection $\mathbf{x}_a \in \mathcal{X}_{k_a}$ contains intrinsic task
features. These describe the work contained in the task instance: payload size,
record count, object count, batch size, partition size, operation mode, or
domain metadata. The projection $\mathbf{c}_a \in \mathcal{C}_{k_a}$ contains
execution-context features. These describe the environment around the attempt:
worker-pool pressure, node pressure, dependency health, network condition, disk
pressure, cache state, retry state, tenant pressure, or domain-specific hidden
state when it is available.

The projections are task-kind-specific because different kinds need different
features. A compute-transform task may project record count and algorithm mode
into $\mathbf{x}_a$, while a replication task may project payload size, replica
count, and divergence level. The same replication task may project peer health,
network condition, disk pressure, and retry state into $\mathbf{c}_a$. A global
feature vector may exist in an implementation, but the mathematical model uses
task-kind-specific spaces to avoid assuming that one fixed metric list explains
all task behavior.

For each task kind $k$, the model may maintain task-kind-specific state
$\theta_k \in \Theta_k$. This state is reusable across attempts. It may contain
dimension descriptors, learned parameters, historical summaries, calibration
values, quantile sketches, residual statistics, priors, or domain rules. It may
be initialized from a manifest, refined from historical observations, and
revised when residuals indicate drift, missing features, or model error.

The roles of the three objects are different:
\[
z_a
\quad \text{is attempt-level input,}
\]
\[
\theta_{k_a}
\quad \text{is reusable task-kind state,}
\]
and
\[
Y_q(a)
\quad \text{is an observed output value, when available.}
\]
This separation prevents leakage between estimation and observation. A value
observed after attempt $a$ may update $\theta_{k_a}$ for future attempts, but it
is not an input feature for estimating the same attempt before it runs.

\subsection{Typed Cost Dimensions}
\label{sec:cost-typed-dimensions}

A cost dimension is not merely a metric name. It is a typed descriptor that
defines the role, unit, attribution, observation, and composition semantics of a
quantity. This is necessary because superficially similar measurements may have
different meanings. For example, elapsed wall-clock time, worker-slot time,
connection hold time, and dependency wait time are all durations, but they do
not describe the same model role.

Let
\[
\mathcal{Q}
=
\mathcal{R}
\dot{\cup}
\mathcal{O}
\]
be the typed disjoint union of resource-demand and worker-occupancy dimensions.
For each $q \in \mathcal{Q}$, define a descriptor
\[
\Delta_q
=
\left(
\rho_q,
u_q,
\mathcal{V}_q,
\alpha_q,
\omega_q,
\kappa_q
\right).
\]
Here $\rho_q$ is the role of the dimension, $u_q$ is the unit,
$\mathcal{V}_q$ is the value domain, $\alpha_q$ is the attribution boundary,
$\omega_q$ is the observation semantics, and $\kappa_q$ is the composition
operator.

Each field prevents a specific modeling error. The role $\rho_q$ prevents
mixing active resource demand with worker occupancy. The unit $u_q$ prevents
combining incompatible quantities, such as bytes and seconds. The value domain
$\mathcal{V}_q$ defines admissible values, usually non-negative values with a
unit. The attribution boundary $\alpha_q$ states when a value belongs to a
particular attempt. The observation semantics $\omega_q$ defines how raw
telemetry becomes an observed value $Y_q(a)$. The composition operator
$\kappa_q$ defines how components combine: by summation, maximum, interval
union, critical path, retry expansion, or another rule.

Several examples illustrate why all fields are needed. For CPU time, the role is
resource demand, the unit is seconds, the attribution boundary identifies CPU
work attributed to the attempt, and composition is usually summation. For memory
peak, the role is resource demand, the unit is bytes, but composition is a
maximum rather than a sum. For worker-slot time, the role is worker occupancy,
the unit is slot-seconds, the attribution boundary is the interval during which
the attempt holds a worker slot, and composition is interval-based. For
connection hold time, the role is worker occupancy, the unit is
connection-seconds, and observation requires acquire and release events for the
connection.

For a task kind $k$, the relevant cost dimensions are
\[
\mathcal{Q}_k
=
\mathcal{R}_k
\dot{\cup}
\mathcal{O}_k
\subseteq
\mathcal{Q}.
\]
The subset $\mathcal{Q}_k$ is part of the task-kind model. A compute transform
may include CPU time, allocation volume, and memory peak. A replication task may
include network output bytes, disk read bytes, disk write bytes, worker-slot
time, connection hold time, and lease time. The model therefore remains
open-world: new domains can add dimensions without requiring one global list of
all possible metrics.

\subsection{Dimension Estimate Objects}
\label{sec:cost-dimension-estimate-objects}

A dimension estimator returns an estimate object rather than only one number.
For a cost dimension $q$, let $\mathcal{E}_q$ be the estimate-object space. The
estimate object for attempt $a$ is
\[
\widehat{E}_q(a) \in \mathcal{E}_q.
\]
This object may contain a point estimate, quantiles, distributional summaries,
conservative envelopes, missingness flags, and other information needed by
diagnostic or policy layers.

The point estimate is a projection:
\[
\widehat{Y}_q(a)
=
\operatorname{point}_q\!\left(\widehat{E}_q(a)\right),
\qquad
\widehat{Y}_q(a) \in \mathcal{V}_q.
\]
The projection $\operatorname{point}_q$ is part of the dimension semantics or
estimator design. It is not necessarily a mean. For some dimensions it may be an
expected value; for others it may be a median, a conservative point, or a value
selected for compatibility with a downstream controller.

When the estimate object contains distributional information, a quantile
projection may also be defined:
\[
\widehat{Y}_{q,p}(a)
=
\operatorname{quantile}_{q,p}\!\left(\widehat{E}_q(a)\right),
\qquad
p \in (0,1).
\]
For example, an estimate object for worker-slot time may contain a median of
two seconds, a 95th percentile of fifteen seconds, and a 99th percentile of two
minutes. A throughput-oriented policy may inspect a point estimate, while a
capacity-protection policy may inspect a high quantile. Both values come from
the same estimate object.

The estimate-object abstraction is needed because queue-worker behavior is often
tail-sensitive. If the model kept only one number per dimension, it would lose
the information needed to distinguish stable short attempts from mostly short
attempts with rare long occupancy tails. The later diagnostic and policy layers
can use the same estimate object without changing the definition of cost.

\subsection{Dimension-Specific Estimators}
\label{sec:cost-dimension-specific-estimators}

The estimator for a task kind is built from dimension-specific estimators. For
each task kind $k$ and dimension $q \in \mathcal{Q}_k$, define
\[
F_{k,q}
:
\mathcal{X}_k
\times
\mathcal{C}_k
\times
\Theta_k
\rightarrow
\mathcal{E}_q.
\]
This signature states that the estimator for dimension $q$ uses intrinsic
features, execution-context features, and reusable task-kind state, and returns
an estimate object for that dimension. It does not require every dimension to
use the same formula internally.

For attempt $a$, the dimension estimate is
\[
\widehat{E}_q(a)
=
F_{k_a,q}\!\left(
\mathbf{x}_a,
\mathbf{c}_a;
\theta_{k_a}
\right).
\]
The full task-kind estimator is the collection of these dimension-specific
estimators:
\[
F_k
=
\left\{
F_{k,q}
\mid
q \in \mathcal{Q}_k
\right\}.
\]
Thus, evaluating the cost model for an attempt means selecting the relevant task
kind, projecting its input features, reading the task-kind model state, and
evaluating the estimators for the dimensions in $\mathcal{Q}_{k_a}$.

Some estimators may be direct rules or learned predictors. Others can be
described through a component model. A common component model separates
intrinsic baseline, contextual amplification, and additive overhead:
\[
\widehat{E}_q(a)
=
\operatorname{estimate}_q\!\left(
\widehat{B}_q(k_a,\mathbf{x}_a;\theta_{k_a}),
\widehat{A}_q(k_a,\mathbf{x}_a,\mathbf{c}_a;\theta_{k_a}),
\widehat{O}_q(k_a,\mathbf{x}_a,\mathbf{c}_a;\theta_{k_a}),
\kappa_q
\right).
\]
The baseline $\widehat{B}_q$ estimates the contribution implied by the task
kind and intrinsic features under a reference context. The amplification factor
$\widehat{A}_q$ estimates how execution context modifies that contribution. The
overhead term $\widehat{O}_q$ estimates additional contribution that is not well
represented as proportional to the baseline. The operator $\kappa_q$ determines
how these components are composed for dimension $q$.

For dimensions whose descriptor permits proportional amplification plus additive
contribution, the point projection may use the affine specialization
\[
\widehat{Y}_q(a)
=
\widehat{B}_q(k_a,\mathbf{x}_a;\theta_{k_a})
\cdot
\widehat{A}_q(k_a,\mathbf{x}_a,\mathbf{c}_a;\theta_{k_a})
+
\widehat{O}_q(k_a,\mathbf{x}_a,\mathbf{c}_a;\theta_{k_a}).
\]
In this form, $\widehat{B}_q$ and $\widehat{O}_q$ have unit $u_q$, while
$\widehat{A}_q$ is dimensionless. Values of $\widehat{A}_q$ above one represent
amplification, values below one represent suppression, and value one represents
no modeled contextual scaling.

The affine form is not a universal law. It is appropriate only when the
dimension supports proportional scaling and additive contribution. CPU time or
some transfer-duration dimensions may fit this form. Memory peak may require a
maximum-based operator. Worker-slot time may require interval or critical-path
semantics. Retry-sensitive work may require retry expansion. A descriptor may
also omit, fix, or replace one of the baseline, amplification, or overhead
components when that component is not meaningful for the dimension.

In the replication example, network output bytes may use payload size as a
baseline feature. Peer health and network condition may amplify connection hold
time or worker-slot time. Retry state may add overhead or expand the effective
number of transfer attempts. These relationships should be encoded in the
dimension estimator or in the Task Behavior Graph that parameterizes it.

\subsection{Execution Overhead and Profiling Overhead}
\label{sec:cost-overhead-types}

The term overhead needs a precise role. Overhead is not a third output family
beside resource demand and worker occupancy. It is a contribution to one or
more existing cost dimensions.

Execution overhead is overhead required by the task execution itself. Examples
include setup, deserialization, connection establishment, lease-extension
bookkeeping, acknowledgement, commit, and cleanup. Depending on the descriptor,
these effects may contribute to CPU time, allocation volume, disk writes,
worker-slot time, connection hold time, or other dimensions.

Profiling overhead is overhead introduced by observation policy. Examples
include tracing, sampling, runtime profiling, high-cardinality label handling,
telemetry export, and additional accounting. Profiling overhead can also
contribute to cost dimensions, but only when the corresponding profiling policy
is enabled.

This distinction motivates two related estimates. A base estimate excludes
additional profiling overhead:
\[
\widehat{\mathbf{C}}_{\mathrm{base}}(a).
\]
A policy-aware estimate includes the overhead of an already selected profiling
policy $\rho$:
\[
\widehat{\mathbf{C}}_{\rho}(a).
\]
The cost model can represent both estimates, but it does not choose $\rho$.
Policy selection is defined later. This separation prevents instrumentation
from being treated as free while keeping the cost model independent of the
policy decision.

\subsection{Observed Values and Residuals}
\label{sec:cost-residuals-usage}

The model distinguishes estimates from observed values. For
$q \in \mathcal{Q}_{k_a}$, the observed value $Y_q(a)$ is not raw telemetry. It
is an attempt-attributed observed value produced according to the descriptor's
observation semantics $\omega_q$ and attribution boundary $\alpha_q$.

For example, node-level CPU utilization is not observed CPU usage for a
particular attempt unless attribution is available. Worker reserve and release
events can produce observed worker-slot time when they delimit the interval
during which the slot is held. Connection acquire and release events can produce
observed connection hold time. Wall-clock duration may provide useful evidence,
but it is not worker-slot time unless the worker slot is held for the whole
interval.

When an observed value exists, the residual is
\[
\epsilon_q(a)
=
Y_q(a)
-
\widehat{Y}_q(a).
\]
A positive residual means the observed value exceeded the point estimate. A
negative residual means the point estimate exceeded the observed value. The
residual is undefined when $Y_q(a)$ is unavailable; absence of a residual does
not imply accuracy.

Residuals are evidence about the relation between the model and the workload.
Large or systematic residuals may indicate missing features, stale context,
hidden dependency state, model misspecification, drift, measurement error, or a
tail event. Residuals can update $\theta_{k_a}$ for future attempts, feed
diagnostics, or trigger a more detailed profiling policy. They are not
themselves cost dimensions.

\subsection{Distributional Cost Estimates}
\label{sec:cost-distributional-estimates}

Attempt-level cost can be variable even within one task kind and one feature
bucket. A single mean-like value may hide behavior that matters operationally.
For queue workers, high-tail worker-slot time can reduce effective worker-pool
capacity; high-tail connection hold time can exhaust dependency pools; retry
storms can produce rare but expensive attempts.

The estimate-object formulation allows each dimension to carry distributional
information. The task-kind model state $\theta_k$ may include empirical
histograms, quantile sketches, parametric distribution parameters, conservative
envelopes, or prediction intervals. The estimate object $\widehat{E}_q(a)$ may
then expose projections such as $\widehat{Y}_{q,0.50}(a)$,
$\widehat{Y}_{q,0.95}(a)$, or $\widehat{Y}_{q,0.99}(a)$.

Distributional estimates are still cost estimates. They should not be confused
with uncertainty or risk. A high 99th percentile is a statement about estimated
cost distribution. Confidence describes how reliable that estimate is. Risk
describes the consequence of underestimating or mishandling it. Later sections
derive diagnostics and risk interpretations from cost estimates, residuals,
missingness, drift, tail ratios, and consequence information.

\subsection{Boundaries of the Cost Model}
\label{sec:cost-model-boundaries}

This section defines attempt-attributed cost: active resource demand and worker
occupancy. It also defines the objects needed to estimate and compare those
quantities: feature projections, model state, typed dimensions, estimate
objects, estimators, observed values, residuals, and distributional projections.

The section does not define the complete role system of the paper. Execution
context, symptoms, outcomes, diagnostics, risk, and policy triggers are not cost
outputs, even though they may influence or consume cost estimates. The next
section defines those non-cost roles as influence domains. Task Behavior Graphs
then use influence domains to connect canonical features, cost estimates,
diagnostics, risk, and profiling-policy triggers in a task-kind-specific model.
