\section{Task Taxonomy}
\label{sec:task-taxonomy}

Task kinds differ in the features that explain their behavior, the cost
dimensions they affect, and the contexts that amplify their costs. This section
introduces a working taxonomy used by later examples. The taxonomy is not a
complete classification of all workloads.

Compute-transform tasks are shaped by record count, payload size, algorithm
mode, CPU pressure, allocation behavior, and memory pressure. Their main cost
dimensions are CPU time, allocation volume, and memory peak.

I/O transfer tasks are shaped by byte count, object count, chunk size, disk
pressure, network condition, and peer health. Their main cost dimensions are
disk reads, disk writes, network input, network output, worker-slot time, and
connection hold time.

External-dependency tasks are shaped by dependency health, fan-out width,
timeout policy, connection-pool pressure, and retry state. Their active CPU
demand may be low while worker-slot time and connection hold time are high.

Replication and repair tasks are shaped by payload size, replica count,
divergence level, peer health, replication lag, disk pressure, and retry state.
They often combine I/O transfer, coordination, and external-dependency behavior.

Coordination-sensitive tasks are shaped by locks, leases, semaphores, tenant
concurrency, and contention. Their active resource demand may be low while
occupancy dominates system impact.

A single task kind may combine several classes. Task Behavior Graphs represent
this compositional behavior through typed nodes and edges rather than by forcing
each task kind into one exclusive class.
