\section{Metrics, Signals, and Features}
\label{sec:metrics-signals-features}

Raw observations are not model variables. Metrics, spans, logs, runtime
profiles, queue events, worker lifecycle events, and domain events become useful
only after they are mapped into typed model variables with assigned influence
domains.

Let $s$ be a raw signal from an observation source. A mapping rule is a typed
transformation
\[
M_j:s\mapsto m_j,
\]
where $m_j$ is a derived model variable and $\delta(m_j)\in\mathcal{D}$ is its
influence-domain assignment. The target may be a canonical input feature, an
observed cost value, a symptom, an outcome, diagnostic evidence, risk evidence,
or a profiling-policy trigger.

The mapping target matters. A packet-loss metric may become a network-condition
context feature. Worker reserve and release events may become observed
worker-slot occupancy. A dependency timeout log may become a symptom if it is
used as intermediate evidence, or an outcome if it is the terminal result of an
attempt. A sustained queue delay may become risk evidence when it describes the
consequence of occupancy or retries.

A single raw observation may map to several variables. For example, a dependency
timeout record may produce a symptom node, an outcome node, and risk evidence.
The one-domain rule still applies: each derived variable receives one primary
domain, and the relations among those variables are represented by graph edges.
This avoids treating one observation as several independent costs.

Canonical features form the input vector $z_a$. Observed output values such as
$Y_q(a)$ are not part of $z_a$ for the same pre-execution estimate; they are
produced during or after execution and used for residuals, calibration,
diagnostics, and future model updates. This separation prevents post-execution
leakage while still allowing the model to learn from historical attempts.
