\section{Task Behavior Manifests}
\label{sec:task-behavior-manifests}

Task Behavior Manifests provide a declarative bridge between the model and
implementation. A manifest declares task classes or task kinds, canonical
features, cost dimensions, observation mappings, graph fragments, diagnostics,
risk rules, and profiling policies.

The manifest layer uses the terminology defined in the previous sections. A
feature declaration must identify whether the variable is intrinsic work or
execution context. A cost-dimension declaration must provide or reference a
descriptor $\Delta_q$. A graph node must have one influence-domain assignment.
A graph edge must respect the permitted domain-flow relation unless the manifest
explicitly marks the edge as implementation-specific and outside the base
behavior model.

A replication manifest may declare \texttt{payload\_size\_bytes} as an
intrinsic feature, \texttt{peer\_health}, \texttt{network\_condition}, and
\texttt{retry\_attempt} as context features, \texttt{network\_output\_bytes}
and \texttt{disk\_read\_bytes} as resource-demand dimensions, and
\texttt{worker\_slot\_time} and \texttt{connection\_hold\_time} as
worker-occupancy dimensions. It may then declare graph edges from payload size
to network output, from peer health to worker-slot time, from retry state to
connection-hold time, from residual diagnostics to retry-storm risk, and from
risk to a tail-aware profiling policy.

Manifests allow domain packs to extend the model without requiring the core
system to know all possible domain metrics in advance. The core requirement is
not that every metric be predefined, but that every introduced variable has a
role, every cost dimension has a descriptor, and every graph edge has a
semantically valid interpretation.
